Robust Control of a Fluxonium Qubit

### 2020-12-10 paper/1introduction.tex
added constraints to opt statement

In this work, we study three robust control techniques that
make the system's state trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{allen2019robust,
    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
    to compute the sensitivity of the quantum state trajectory
    to deviant parameters.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular, taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation type decoherence and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

>>>>

In this work, we study three robust control techniques that
make the system's state trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{allen2019robust,
    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
    to compute the sensitivity of the quantum state trajectory
    to deviant parameters.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits: mitigating
longitudinal relaxation type decoherence
by taking advantage of the $T_{1}$-dependence of the controls
and performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

### 2020-12-10 paper/2background.tex
added constraints to opt statement

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
$u_k \in \R^m$ are the augmented state and input augmented control variables,
$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
$h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

Many techniques have been proposed for solving \eqref{opt:discrete_trajopt}. Standard 
methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
have combined principles from both of these approaches.

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^{l + m} \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR,
the Lagrange multiplier estimates are updated according to,
\begin{equation}
	\lambda \gets \lambda + \mu c(z) ,
\end{equation}
the penalty terms are updated, and the process repeats until convergence.

@@

\todo{Ask Dave to review paragraph}
As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
for imposing arbitrary, simultaneous constraints
on both the state and the controls and does not restrict
optimization to the constraint manifold. First-order solvers like GRAPE
rely on unconstrained optimization to satisfy desiderata on the state,
such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
and smoothing the controls below a threshold. The unconstrained approach relies
on a careful choice of competing penalty weights, which is cumbersome and often intractable.
To impose constraints on the controls, such as amplitude restrictions, first-order solvers
employ projected gradient descent methods which restrict optimization to the constraint manifold and
may hinder convergence to the optimal solution \cite{clarkson2010coresets,
  hauswirth2016projected}.
The ALTRO solver delays projecting onto the constraint manifold until the ALM
solve is already at a coarse tolerance, achieving machine-precision satisfaction
for constraints on the controls and the state. These advantages are critical
for this work, where multiple constraints must be satisfied in addition
to minimizing auxiliary metrics.

>>>>

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini!}[2]
    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k)}{}{} \label{eq:gcostfun}
    \addConstraint{x_{k+1} = f(x_k,u_k)} \label{eq:gdyncon}
    \addConstraint{g_k(x_k,u_k)}{\leq 0} \label{eq:ineqcon}
    \addConstraint{h_k(x_k,u_k)}{=0} \label{eq:eqcon}
\end{mini!}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
$u_k \in \R^m$ are the state and control variables,
$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
$h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

Many techniques have been proposed for solving \eqref{eq:gcostfun}--\eqref{eq:eqcon}. Standard 
methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
have combined principles from both of these approaches.

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^{l + m} \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
of penalty weights, $\mu$, whose magnitudes depend on whether
the corresponding constraint is active or inactive.
For ALTRO, $f(z)$ is \eqref{eq:gcostfun}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR,
the Lagrange multiplier estimates are updated according to,
\begin{equation}
	\lambda \gets \lambda + \mu c(z)
\end{equation}
the penalty terms are updated, and the process repeats until convergence.

@@

As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
for imposing arbitrary, simultaneous constraints
on both the states and the controls that converge superlinearly to tight tolerances. 
First-order solvers such as GRAPE
use unconstrained optimization to satisfy desiderata on the states,
such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
and smoothing the controls below a threshold. The unconstrained approach relies
on a careful choice of competing penalty weights, which is cumbersome and often intractable.
To impose constraints on the controls, such as amplitude restrictions, first-order solvers
employ projected gradient descent methods which restrict
optimization to the constraint manifold and
may hinder convergence to the optimal solution \cite{clarkson2010coresets,
  hauswirth2016projected}.
Conversely,
the ALTRO solver delays projecting onto the constraint manifold defined by
\eqref{eq:ineqcon} and \eqref{eq:eqcon} until the ALM
solve is already at a coarse tolerance, achieving machine-precision satisfaction
for constraints on the controls and the states. These advantages are critical
for this work, where multiple constraints must be satisfied in addition
to minimizing auxiliary metrics.

### 2020-12-10 paper/3fluxonium.tex
added constraints to opt statement

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1 \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
The augmented state and augmented controls are:
\begin{equation}
  x = \begin{bmatrix} \psi^{1} \\ \psi^{2} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial states $\psi^{1}_{1}$, $\psi^{2}_{1}$
are basis vectors of the Hilbert space $\ket{0}, \ket{1}$.
The final states $\psi^{1}_{f}$, $\psi^{2}_{f}$ are the image
of the initial states under the desired gate.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
given in \cite{leung2017speedup},
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
  \label{eq:isomorphism}
\end{equation}
The discrete dynamics function \eqref{eq:dyn_con} integrates the states
using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
and integrates the moments of the flux amplitude.
Exposing lower order moments of the flux amplitude allows us to penalize their norms
in \eqref{eq:costfun}, smoothing the flux amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
Including objectives at each knot point smoothens the optimization landscape, though
the importance of the final augmented state objective is encoded in the relative
weight $Q_{N} \sim N \cdot Q_{k}$.
We choose $Q$ to be diagonal because it is computationally efficient. This
metric penalizes phase differences between the states, although
phase-insensitive metrics such as infidelity may be employed.

We impose constraints that
reflect the physical limitations of the apparatus and 
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a_{1} = a_{N} = 0$.
Furthermore, we require $\int a_{N} \ dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a_{k} \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require
$\psi_{N} = \psi_{f}$.
The zero net flux and final state constraints
are satisfied if the final augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition, the norm of the states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
of subsets from $x_{k}$ is implied in
the constraint equations.

>>>>

The optimization problem takes the form:
\begin{mini!}[2] 
  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
  \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}} \label{eq:tstate_con}
  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
      \rvert}^{2} = 1} \label{eq:statenorm_con}
  \addConstraint{{\textstyle \int^{t_{1}}_{0} a \ \mathrm{d}t} = \pdv*{a_{1}}{t} = 0}
  \label{eq:icintd_con}
  \addConstraint{{\textstyle \int^{t_{N}}_{0}} a \ \mathrm{d}t= 0} \label{eq:znf_con}
  \addConstraint{a_{1} = a_{N} = 0} \label{eq:concat_con}
  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz}} \label{eq:amp_con}
\end{mini!}
The augmented state and augmented controls are:
\begin{equation}
  x = \begin{bmatrix} \psi^{0} \\ \psi^{1} \\ \int^{t}_{0} a \ \mathrm{d}t\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
Including objectives at each knot point smoothens the optimization landscape, though
the importance of the final objective is encoded in the relative
weight $Q_{N} \sim N \cdot Q_{k}$.
We choose $Q$ to be diagonal because it is computationally efficient. This
metric penalizes phase differences between the states, although
phase-insensitive metrics such as infidelity may be employed.

We impose constraints that
reflect the physical limitations of the apparatus and 
improve the experimental realization of the control pulse.
The discrete dynamics \eqref{eq:dyn_con} integrates the states
according to the TDSE \eqref{eq:tdse} and the
fluxonium Hamiltonian \eqref{eq:hamiltonian}
and integrates the moments of the flux amplitude.
Exposing lower order moments of the flux amplitude allows us to penalize their norms
in \eqref{eq:costfun}, smoothing the flux amplitude
and mitigating AWG ringing due to high frequency transitions.
We ensure the desired gate is achieved by imposing the constraints
\eqref{eq:istate_con} and \eqref{eq:tstate_con}.
The initial states in \eqref{eq:istate_con}
are chosen to span the Hilbert space.
The final states in \eqref{eq:tstate_con} are the image
of the initial states under the desired gate.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
given in \cite{leung2017speedup},
\begin{equation}
  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
  \label{eq:isomorphism}
\end{equation}
We ensure the solver does not take advantage of discretization
error by imposing the normalization constraint \eqref{eq:statenorm_con}.
The constraint \eqref{eq:icintd_con} is an initial condition.
We ensure the pulse has zero net flux by imposing \eqref{eq:znf_con},
which mitigates the hysteresis ubiquitous in flux bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions by imposing
the boundary condition \eqref{eq:concat_con}.
Finally, we ensure the two-level
approximation \eqref{eq:hamiltonian} remains valid
by imposing the amplitude constraint \eqref{eq:amp_con}.
All of the gates presented in this work achieve a maximum
individual constraint violation of $10^{-8}$
for \eqref{eq:dyn_con}--\eqref{eq:amp_con}.

### 2020-12-10 paper/5static.tex
Brian edits
added constraints to opt statement

\todo{Ask Brian to review this paragraph and next two.
Are the optimization setups clear?}
The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
We append four sets of two sample states $\psi^{\pm}$
to the augmented state vector \eqref{eq:astatecontrols}.
We initialize each pair of samples in the same state.
The four initial states are chosen so that their outerproducts
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. The infidelity between
the samples and their final states are penalized
by adding a term
$\sum_{k=1}^{N} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
to the objective \eqref{eq:costfun}.

>>>>

The sampling method simultaneously optimizes over copies of a state,
each of which evolves with a distinct deviant parameter value.
We append four sets of two sample states $\psi^{\pm}$
to the augmented state vector \eqref{eq:astatecontrols}.
We initialize each pair of samples in the same state.
The four initial states are chosen so that their outerproducts
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
The discrete dynamics \eqref{eq:dyn_con} is modified so that the
samples evolve under the Hamiltonian given in \eqref{eq:hamiltonian}
with $f_{q} \gets f_{q} \pm \sigma_{f_{q}}$ for a fixed $\sigma_{f_{q}}$.
The final state
for each sample is the image of its initial state
under the desired gate. The infidelity between
the samples and their final states are penalized
by adding a term
$\sum_{k=1}^{N} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
to the objective \eqref{eq:costfun}.

### 2020-12-10 paper/a3derivative.tex
added constraints to opt statement

The dynamics for the derivative methods can be integrated efficiently using exponential integrators
\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
General exponential integrators break the dynamics into a linear term and a non-linear term.
For example, consider
integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact solution is:
\begin{equation}
  \label{eq:dgeneralexp}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k})
    \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
    N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
method \cite{berland2005solving},
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Integration accuracy is not of the utmost importance because the
state derivatives guide the optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

>>>>

The dynamics for the derivative methods can be integrated efficiently using exponential integrators
\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
General exponential integrators break the dynamics into a linear term and a non-linear term.
For example, the dynamics of the first state derivative in
units of $i\hbar = 1$ are $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact solution is:
\begin{equation}
  \label{eq:dgeneralexp}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k})
    \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
    N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
method \cite{berland2005solving},
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Integration accuracy is not of the utmost importance because the
state derivatives guide the optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

### 2020-12-10 paper/a4time.tex
added constraints to opt statement

\section{Computational Performance \label{appendix:time}}
We provide runtimes for our optimizations and comment on the scaling of the
robustness methods. The runtimes for
the base optimization in Section \ref{sec:fluxonium},
the longitudinally aware optimization in Section \ref{sec:longitude},
and the robust optimizations in Section \ref{sec:static}
are presented in Table \ref{tab:time}
for a $Z/2$ gate at multiples of the analytic $Z/2$ gate time.
Compiler optimizations for statically-sized arrays were utilized
for all methods except for the unscented sampling method.
The unscented sampling method's augmented state vector size
was too large to take advantage of these optimizations,
adversely affecting its run time.
For the robust methods, the runtime scales super-linearly with the
gate time. We expect the runtime to scale linearly with the gate time
for large gate times because
the number of knot points scales linearly with the gate time.
We observe this behavior for the base method, which has a smaller
augmented state vector size than the robust methods.
We performed optimizations on a single CPU thread in this work as a proof
of concept. Future work will parallelize the robustness
methods across the initial states in the operator basis and utilize GPUs
\cite{leung2017speedup},
allowing for fast optimization on large Hilbert spaces.
\vspace*{\intextsep}

\begin{table}[H]
  \begin{tabular} {c | c | c | c | c | c | c}
    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
    \hline
    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
    \hline
    $18$ & 0.155 & 1.688 & 1.773 & 210.573 & 16.713 & 48.398\\ % 0.155049
    $36$ & 7.014 & - & 48.213 & 4566.236 & 67.838 & 81.030\\
    $72$ & 15.906 & - & 281.372 & 16575.308 & 266.997 & 332.182\\
  \end{tabular}
  \caption{
    Runtimes for $Z/2$ optimizations.
    Programs were executed on a single thread of an
    AMD Ryzen Threadripper 3970X 32-Core Processor.
  }
  \label{tab:time}
\end{table}

>>>>

\section{Computational Performance \label{appendix:time}}
We provide runtimes for our optimizations and comment on the scaling of the
robustness methods. The runtimes for
the base optimization in Section \ref{sec:fluxonium},
the longitudinally aware optimization in Section \ref{sec:longitude},
and the robust optimizations in Section \ref{sec:static}
are presented in Table \ref{tab:time}
for a $Z/2$ gate at multiples of the analytic $Z/2$ gate time.
Compiler optimizations for statically-sized arrays were utilized
for all methods except for the unscented sampling method.
The unscented sampling method's augmented state vector size
was too large to take advantage of these optimizations,
which adversely affected its run time.
The runtimes of the robust methods scale
super-linearly with the gate time.
For large gate times, we expect the runtimes
to scale linearly with the gate time
because the number of knot points scales linearly with the gate time.
We observe this trend for the base method, which has a smaller
augmented state vector size than the robust methods.
We performed optimizations on a single CPU thread in this work as a proof
of concept. Future work will parallelize the robustness
methods across the initial states in the operator basis and utilize GPUs
\cite{leung2017speedup},
which will enable fast optimizations on large Hilbert spaces.
\vspace*{\intextsep}
\begin{table}[H]
  \begin{tabular} {c | c | c | c | c | c | c}
    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
    \hline
    $t_{N}$ (ns) & Base & Long. & S & SU & D1 & D2\\
    \hline
    $18$ & 0.155 & 1.688 & 1.773 & 210.573 & 16.713 & 48.398\\ % 0.155049
    $36$ & 7.014 & - & 48.213 & 4566.236 & 67.838 & 81.030\\
    $72$ & 15.906 & - & 281.372 & 16575.308 & 266.997 & 332.182\\
  \end{tabular}
  \caption{
    Runtimes for $Z/2$ optimizations.
    Programs were executed on a single thread of an
    AMD Ryzen Threadripper 3970X 32-Core Processor.
  }
  \label{tab:time}
\end{table}
